---
title: "NHIS 2021 Data Analysis Group Project 2 (Group 5)"
author: "Karinna Alatorre, Celia Lopez, Bi Na, and Alyssa Tagle"
date: "2026-04-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#Day 1: Project Setup & Data Import
#Task 2: Environment Setup & Data Import

```{r}
# Load packages
library(tidyverse)

# Load data
file_path <- "NHIS_Data_2021.csv"

if (!file.exists(file_path)) {
  stop("Dataset not found. Please put 'NHIS_Data_2021.csv' in the same folder.")
}

NHIS_Data_2021 <- read_csv(file_path)

# Explore data
glimpse(NHIS_Data_2021)
summary(NHIS_Data_2021)
head(NHIS_Data_2021)
```

#Day 2: Data Cleaning & Management
#Task 1&2: Variable Selection & Missing Value Check & Recoding Variables
```{r}
nhis_subset <- NHIS_Data_2021 %>%
  select(
    AGEP_A,
    WEIGHTLBTC_A,
    HEIGHTTC_A,
    SEX_A,
    HISPALLP_A,
    EDUCP_A,
    PHSTAT_A,
    LSATIS4R_A
  )

nhis_clean <- nhis_subset %>%
  filter(
    !AGEP_A %in% c(97, 98, 99),
    !WEIGHTLBTC_A %in% c(996, 997, 998, 999),
    !HEIGHTTC_A %in% c(96, 97, 98, 99),
    !SEX_A %in% c(7, 9),
    !HISPALLP_A %in% c(97, 98, 99),
    !EDUCP_A %in% c(97, 98, 99),
    !PHSTAT_A %in% c(7, 8, 9),
    !LSATIS4R_A %in% c(7, 8, 9)
  )

# Check missing values
anyNA(nhis_clean)
dim(nhis_clean)
```

#Task 2: Recoding Variables
```{r}
nhis_clean <- nhis_clean %>%
  mutate(
    EDUCP_A = case_when(
      EDUCP_A %in% 1:4 ~ "Less than High School",
      EDUCP_A == 4 ~ "High School Graduate",
      EDUCP_A %in% 5:7 ~ "Some College Education",
      EDUCP_A %in% 8:10 ~ "College Graduate or better",
      TRUE ~ NA_character_
    ),
    EDUCP_A = factor(
      EDUCP_A,
      levels = c(
        "Less than High School",
        "High School Graduate",
        "Some College Education",
        "College Graduate or better"
      )
    )
  )

# Final checks
summary(nhis_clean)
dim(nhis_clean)
View(nhis_clean)

